Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-02-16 grove-paper.tex
make lfp a math operator
move grove example to right of its graph

% graph decomposition operators
\DeclareMathOperator{\expr}{\text{expr}}
\DeclareMathOperator{\patt}{\text{patt}}
\DeclareMathOperator{\type}{\text{type}}
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\ancestors}{\text{ancestors}}
\DeclareMathOperator{\vertices}{\text{vertices}}
\DeclareMathOperator{\children}{\text{children}}

@@

\newvertex{DecompPlusA}{+}
\newedge{DecompTimesA}{L}
\newvertex{DecompTimesA}{*}
\newedge{DecompPlusAX}{R}
\newvertex{DecompX}{x}
\newedge{DecompTimesAX}{R}
%
\newedge{DecompPlusB}{L}
\newvertex{DecompPlusB}{+}
\newedge{DecompTimesB}{R}
\newvertex{DecompTimesB}{*}
\newedge{DecompY}{R}
\newvertex{DecompY}{y}
\newedge{DecompZ}{R}
\newvertex{DecompZ}{z}
%
\newedge{DecompTimes}{L}
%
\newcommand{\figureDecompExample}{
\begin{figure}
  \centering
  \begin{tikzpicture}
    \path (0cm,0cm) graph[graph style] {
      root[root style] -> {
       a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
        "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
         {},
         b/"\vDecompX" [> "\eDecompTimesAX"']
        },
        {}
       }
      },
    };
    \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
    %
    \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
    \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
    \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
    \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
    \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
    \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
    %
    \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
  \end{tikzpicture}
  \caption{Caption goes here}%
  \label{fig:Decomposition example}
\end{figure}
}

@@

A grove is a 4-tuple of the following form:
\[
  (e_r, MP, NP, W)
\]
where
\begin{align*}
  e_r &= \expr(v_0) \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example} is
\begin{align*}
  e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
  MP &= \{x^{\id{42}}\} \\
  NP &= \{z^{\id{52}}\} \\
  W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
\end{align*}

@@

% TODO: make macros for + and *
% TODO: comment these out:
%   % | Exp_nil
%   % | Exp_cons
%   % | Exp_case
%   % | Typ_list

@@

  \begin{align*}
    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
    \ancestors(v) &= \left( \textrm{lfp}(\ancestors') \right)(v) \\
    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
      \end{align*}
  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of $a$.}
  \label{fig:Decomposition expr}
\end{figure}

>>>>

% graph decomposition operators
\DeclareMathOperator{\expr}{\text{expr}}
\DeclareMathOperator{\patt}{\text{patt}}
\DeclareMathOperator{\type}{\text{type}}
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\ancestors}{\text{ancestors}}
\DeclareMathOperator{\vertices}{\text{vertices}}
\DeclareMathOperator{\children}{\text{children}}
\DeclareMathOperator{\lfp}{\text{lfp}}

@@

\newvertex{DecompPlusA}{+}
\newedge{DecompTimesA}{L}
\newvertex{DecompTimesA}{*}
\newedge{DecompPlusAX}{R}
\newvertex{DecompX}{x}
\newedge{DecompTimesAX}{R}
%
\newedge{DecompPlusB}{L}
\newvertex{DecompPlusB}{+}
\newedge{DecompTimesB}{R}
\newvertex{DecompTimesB}{*}
\newedge{DecompY}{R}
\newvertex{DecompY}{y}
\newedge{DecompZ}{R}
\newvertex{DecompZ}{z}
%
\newedge{DecompTimes}{L}
%
\newcommand{\figureDecompExample}{
\begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
    \centering
  \begin{tikzpicture}
    \path (0cm,0cm) graph[graph style] {
      root[root style] -> {
       a/{\vDecompPlusA} [> "\eSimpleTimes"] -> {
        "\vDecompTimesA" [> "\eDecompTimesA"'] -> {
         {},
         b/"\vDecompX" [> "\eDecompTimesAX"']
        },
        {}
       }
      },
    };
    \path [draw,->] (a) to ["\eDecompPlusAX",out=-60,in=60] (b);
    %
    \path (2cm,-0.85cm) graph[graph style] { plus/"\vDecompPlusB" };
    \path (3cm,-0.85cm) graph[graph style] { times/"\vDecompTimesB" };
    \path (3.25cm,-1.75cm) graph[graph style] { y/"\vDecompY" };
    \path[draw,->] (times) to ["\eDecompPlusB"'anchor=north,out=-135,in=-45] (plus);
    \path[draw,->] (plus) to ["\eDecompTimesB"anchor=south,out=45,in=135] (times);
    \path[draw,->] (times) to ["\eDecompY"anchor=west] (y);
    %
    \path (1.55cm,-2.1cm) graph[graph style] { plus/"\vDecompZ" };
  \end{tikzpicture}
    \caption{subcaption 1 goes here}%
    \label{fig:Decomposition example graph}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{align*}
      e_r &= \hole~\texttt{*}^{\id{40}}~\multiVertex{42}~\texttt{+}^{\id{38}}~\multiVertex{42} \\
      MP &= \{x^{\id{42}}\} \\
      NP &= \{z^{\id{52}}\} \\
      W &= \{\hole~\texttt{+}^{\id{46}}~\cycleVertex{46}~\texttt{*}^{\id{48}}~y^{\id{50}}\} \\
    \end{align*}
    \caption{subcaption 2 goes here}%
    \label{fig:Decomposition example grove}
  \end{subfigure}
  \caption{Main caption goes here}%
  \label{fig:Decomposition example}
\end{figure}
}

@@

A grove is a 4-tuple of the following form:
\[
  (e_r, MP, NP, W)
\]
where
\begin{align*}
  e_r &= \expr(v_0) \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example graph} is
given in \autoref{fig:Decomposition example grove}.

@@

% TODO: comment these out:
%   % | Exp_nil
%   % | Exp_cons
%   % | Exp_case
%   % | Typ_list

@@

  \begin{align*}
    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
    \ancestors(v) &= \left( \lfp(\ancestors') \right)(v) \\
    \ancestors'(v) &= \parents(v) \cup \ancestors'(\parents(v)) \\
    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
      \end{align*}
  \caption{Caption2 goes here. Note that $\ancestors$ is defined as the least fixed point of $\ancestors'$.}
  \label{fig:Decomposition expr}
\end{figure}
